A formalization of semi-simplicial and semi-cubical sets in dependent type theory

### 2023-01-13 paper/paper.tex
Update

\begin{document}
\maketitle
\begin{abstract}
  Constructions such as semi-simplicial and semi-cubical sets can be defined in the ``usual way'' as presheaves over respectively the semi-simplex or semi-cube category, which we shall call ``fibered'' definitions, but also, as done e.g. in Voevodsky~\cite{} and in previous work~\cite{} as an ``dependently-typed'', or, say, ``indexed'' construction.

@@

\subsection{Working with sets in type theory}
Martin-Löf's Type theory~\cite{martinlof84,?} is a logical formalism based on the notion of $\Type$ rather than \Set. It is the core of several tools for the formalization of mathematics such as Coq~\cite{} or Lean~\cite{}. Type Theory is a flexible formalism supporting different models. One of this model interprets types as topological spaces, where equality is interpreted as path, and substitutivity of equality as transport, leading to the development of Homotopy Type Theory~\cite{hottbook}, a foundation of mathematics alternative to set theory. Conversely, a notion of sets can be recovered within Type Theory as \U, denoting the subtype of $\Type$ for which paths are degenerated, what can be expressed by the property of Uniqueness of Identity Proofs (UIP). Technically, this is expressed as a structure equipping a domain $\Dom$ with the property $\UIP$:

\begin{align*}
  \Dom & : \Type                                          \\
  \UIP & : \forall x y : \Dom, \forall p q : x = y, p = q
\end{align*}

In the $\U$ universe, the following properties hold:

@@

@@

\begin{equation*}
  \Sigma S: \U\;.\;S\rightarrow T \simeq T \rightarrow \U
\end{equation*}

@@

We can iteratively apply the above correspondence to obtain an indexed representation of cubical sets.

>>>>

\begin{document}
\maketitle
\begin{abstract}
  Constructions such as semi-simplicial and semi-cubical sets can be defined in the ``usual way'' as presheaves over respectively the semi-simplex or semi-cube category, which we shall call \emph{fibered} definitions, but also, as done e.g. in Voevodsky~\cite{voevodsky12} and in previous work~\cite{herbelin15} as a dependently-typed, or, say, \emph{indexed} construction.

@@

\subsection{Working with sets in type theory}
Martin-Löf's Type theory~\cite{martinlof75,martinlof84} is a logical formalism based on the notion of a \emph{type} rather than that of a \emph{set}. It can be seen as a foundation of mathematics alternative to set theory and is the core of several tools for the formalization of mathematics such as Coq~\cite{coq22} or Lean~\cite{lean15}. Type theory is a flexible formalism supporting different models. One of this model interprets types as topological spaces, where equality is interpreted as path and substitutivity of equality as transport. This model supports the univalence principle stating that equality of types mimics equivalence of types, leading to the development of Homotopy Type Theory~\cite{hottbook}. In type theory, types are organized in a hierarchy of universes written $\Type_m$ for $m$ a natural number. Typical types in type theory are the types of dependent pairs, called $\Sigma$-types, with projections $\hd$ and $\tl$, the types of dependent functions, called $\Pi$-types, a distinguished singleton type called $\textsf{unit}$ type, the type of Boolean values, called $\textsf{bool}$, the type of natural numbers.

A notion of sets can be recovered in each universe as $\U[m]$, denoting the subtype of $\Type_m$ for which paths are degenerated, what can be expressed by the property of Uniqueness of Identity Proofs (UIP). Technically, this is expressed as a structure equipping a domain $\Dom$ with the property $\UIP$:

\begin{align*}
  \Dom & : \Type_m                                          \\
  \UIP & : \forall x y : \Dom, \forall p q : x = y, p = q
\end{align*}

In $\U[m]$, the following properties hold:

@@

Unless otherwise specified, we fix a universe level $m$ and abbreviate
$\Type_m$ as $\Type$ and $\U[m]$ as $\U$.

@@

\begin{equation*}
  \Sigma S: \U\;.\;S\rightarrow T ~\simeq~ T \rightarrow \U
\end{equation*}

@@

We can iteratively apply the above correspondence to obtain an indexed representation of semi-cubical sets.
